### 2022-07-11 manuscript/supplement.tex
minor updates + backing up manuscript draft

\documentclass[letterpaper,man,natbib]{apa6}  %Leave this stuff alone

\usepackage[english]{babel}
\usepackage[utf8x]{inputenc}
\usepackage{amsmath}
\usepackage{graphicx}
\usepackage{booktabs}
\usepackage[export]{adjustbox}

% dashed line (https://tex.stackexchange.com/questions/112343/beautiful-table-samples)
\usepackage{array}
\usepackage{arydshln}
\setlength\dashlinedash{0.2pt}
\setlength\dashlinegap{1.5pt}
\setlength\arrayrulewidth{0.3pt}

\title{Supplemental materials for "Bifactor models of the ``Maltreatment and Abuse Chronology of Exposure' (MACE) Scale''}
\shorttitle{Bifactor models of MACE (Suppelment)}
\author{Samuel Zorowitz$^1$, Lauri Tuominen$^{2}$}
\affiliation{$^1$Princeton Neuroscience Institute, Princeton University, USA\\$^2$The Royal’s Institute of Mental Health Research, University of Ottawa, Canada}

\begin{document}
\maketitle

\begin{table}
\begin{tabular*}{\textwidth}{clccc}
\centering
& & \multicolumn{3}{c}{Tetrachoric Corr.} \\
\cmidrule(lr){3-5}
\# & Item & Pair & 1 & 2 \\
\toprule
6 & {\small Intentionally pushed, grabbed, shoved, slapped, pinched, punched} & \\
  & {\small or kicked you.} & \\
7 & {\small Hit you so hard that it left marks for more than a few minutes.} & (7,8) & 0.774 & 0.763 \\
8 & {\small Hit you so hard, or intentionally harmed you in some way, that} & \\
& {\small you received or should have received medical attention.} & \\
\midrule
9	& {\small Spanked you on your buttocks, arms or legs.} & (9,10) & 0.977 & 0.789 \\
10	& {\small Spanked you on your bare (unclothed) buttocks.} & (9,11) & 0.980 & 0.757 \\
11 & {\small Spanked you with an object such as a strap, belt, brush, paddle,} & (10,11) & 0.474 & 0.615 \\
 & {\small rod, etc.} & \\
\midrule
13 & {\small Touched or fondled your body in a sexual way.} & (13,14) & 0.949 & 0.961 \\
14 & {\small Had you touch their body in a sexual way.} & \\
\midrule
15 & {\small Hit your sibling (stepsibling) so hard that it left marks for more} & (15,16) & 0.961 & 0.976 \\
& {\small than a few minutes.} & \\
16 & {\small Hit your sibling (stepsibling) so hard, or intentionally harmed him} & \\
& {\small or her in some way, that he/she received or should have received} & \\
& {\small medical attention.} & \\
\midrule
19 & {\small Had you touch their body in a sexual way.} & (19,20) & 0.920 & 0.916 \\
20 & {\small Actually had sexual intercourse (oral, anal or vaginal) with you.} & \\
\midrule
21 & {\small Saw adults living in the household push, grab, slap or throw} & (21,22) & 0.920 & 0.917 \\
& {\small something at your mother (stepmother, grandmother).} & \\
22 & {\small Saw adults living in the household hit your mother, stepmother, or} & (21,23) & 0.831 & 0.775 \\
& {\small grandmother so hard that it left marks for more than a few minutes.} & \\
23 & {\small Saw adults living in the household hit your mother, stepmother, or} & (22,23) & 0.948 & 0.916 \\
& {\small grandmother so hard, or intentionally harm her in some way, that} & \\
& {\small she received or should have received medical attention.} & \\
\midrule
\end{tabular*}
\end{table}

\begin{table}[ht!]
\begin{tabular}{clccc}
\centering
24 & {\small Saw adults living in the household push, grab, slap or throw} & (24,25) & 0.991 & 0.824 \\
& {\small something at your father (stepfather, grandfather).} & \\
25 & {\small Saw adults living in the household hit your father, stepfather, or} & \\
& {\small grandfather so hard that it left marks for more than a few minutes.} & \\
\midrule
33 & {\small Intentionally pushed, grabbed, shoved, slapped, pinched, punched,} & (33, 34) & 0.833 & 0.833 \\
& {\small or kicked you.} & \\
34 & {\small Hit you so hard that it left marks for more than a few minutes.} & (33, 35) & 0.677 & 0.795 \\
35 & {\small Hit you so hard, or intentionally harmed you in some way, that} & (34, 35) & 0.902 & 0.915 \\
& {\small you received or should have received medical attention.} & \\
\midrule
36 & {\small Forced you to engage in sexual activity against your will.} & (36,37) & 0.980 & 0.967 \\
37 & {\small Forced you to do things sexually that you did not want to do.} & \\
\bottomrule
\end{tabular}
\end{table}

\end{document}